---
title: "hw04_Tidy_data_and_Joins"
author: "Xuemeng Li"
date: "06/10/2019"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(gapminder)
library(tidyverse)
```

# Overview
The goal of this homework is to solidify data wrangling skills by working some realisic problems in the grey area between data aggreagation and data reshaping

# Exercise 1: Univariate Data Reshaping - Univraiate Option 1

## 1. Data in wider format 
Make a tibble with one row per year, and columns for life expectancy for two or more countries. Countries are selected for canada, China and Afghanistan
```{r}
LifeExp1 <- gapminder %>%
  select(year, lifeExp, country) %>%
  filter(country == "Canada" | country == "China" | country == "Afghanistan") %>%
  pivot_wider (id_col = year,
               names_from = country,
               values_from = lifeExp
                 )
DT::datatable(LifeExp1)
```


## 2. Plot of the wide data
Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.
```{r}
LifeExp1 %>%
  ggplot (aes(x=China, y=Canada)) +
  geom_point(color = "red", size = 2) +
  labs (title = "Life expectancy of Canada vs China in 1952-2007", x = "China (Age/Years)", y = "Canada (Age/Years)")
  

  
```


## 3. Re-lengthening the wider data
```{r}
LifeExp1 %>%
  pivot_longer(col = -year,
              names_to = "Countries",
              values_to = "lifeExp"
               ) %>%
  DT::datatable()
```


# Exercise 2: Multivariate Data Reshaping - Multivariate Option 1
## 1. Putting data in wider format
Make a tibble with one row per year, and columns for life expectancy and GDP per capita (or two other numeric variables) for two or more countries.
```{r}
LifeExp2 <- gapminder %>%
  select (year, lifeExp, gdpPercap, country) %>%
  filter(country == "Canada" | country == "China" | country == "Afghanistan") %>%
  pivot_wider (id_cols = c(year,country),
              names_from = "country",
              names_sep = " ",
              values_from = c(gdpPercap, lifeExp)
              )

DT::datatable(LifeExp2)
```

## 2. Re-lengthening the data

```{r}
LifeExp2 %>%
  pivot_longer (cols = -year,
                names_sep = " ",
                names_to = c(".value", "country")
                  )%>%
  DT::datatable()
```

# Exercise 3: Table Joins

```{r}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```

## Practice 1
**For each guest in the guestlist (guest tibble), add a column for email address, which can be found in the email tibble.**

```{r}
guest
```

```{r}
email
```


```{r}
emailList <- email %>%
  separate_rows(guest, sep = ", ")

guest %>%
  left_join (emailList, by = c("name" = "guest")) %>%
  DT::datatable()
```


## Practice 2
**3.2 Who do we have emails for, yet are not on the guestlist?**

```{r}
emailList <- email %>%
  separate_rows(guest, sep = ", ")

guest %>%
  anti_join(emailList, by = c("name" = "guest")) %>%
  select (name) %>%
  DT::datatable()
```

## Practice 3
**3.3 Make a guestlist that includes everyone we have emails for (in addition to those on the original guestlist).**

```{r}
guest %>%
  full_join(emailList, by = c("name" = "guest")) %>%
  select (name, email) %>%
  DT::datatable()
```