---
title: "hw02_ggplot2_dplyr"
author: "Xuemeng Li"
date: "27/09/2019"
output: 
  html_document:
    toc: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Objective
The goal is to get feet wet with ggplot2 and dplyr to explore a dataset.

```{r}
library(gapminder)
library(tidyverse)
```

# Exercise 1: Basic dplyr

##1.1 Use filter() to subset the gapminder data to three countries of your choice in the 1970’s.
```{r}
gapminder %>%
  filter (country == "China" | country == "Canada" | country == "Austria") %>%
  filter (year >= 1970 & year <1980)

```

##1.2 Use the pipe operator %>% to select “country” and “gdpPercap” from your filtered dataset in 1.1.
```{r}
gapminder %>%
  filter (country == "China" | country == "Canada" | country == "Austria") %>%
  filter (year >= 1970 & year <1980) %>%
  select (country, gdpPercap)
```

##1.3 Filter gapminder to all entries that have experienced a drop in life expectancy. Be sure to include a new variable that’s the increase in life expectancy in your tibble. Hint: you might find the lag() or diff() functions useful.
```{r}
gapminder %>%
  group_by(country) %>%
  arrange(year)%>%
  mutate (inc_life_exp = lifeExp - lag(lifeExp)) %>%
  filter (inc_life_exp < 0) %>%
  DT::datatable()
```

##1.4 Filter gapminder so that it shows the max GDP per capita experienced by each country. Hint: you might find the max() function useful here.
```{r}
gapminder %>%
  select (country, year, gdpPercap) %>%
  group_by(country) %>%
  mutate (max_gdpPercap = max(gdpPercap)) %>%
  filter (max_gdpPercap == gdpPercap) %>%
  DT::datatable()
```

##1.5 Produce a scatterplot of Canada’s life expectancy vs. GDP per capita using ggplot2, without defining a new variable. That is, after filtering the gapminder data set, pipe it directly into the ggplot() function. Ensure GDP per capita is on a log scale.
```{r}
gapminder %>%
  filter (country == "Canada") %>%
  ggplot (aes(gdpPercap, lifeExp)) +
  geom_point () +
  scale_x_log10()
  
```

# Exercise 2: Explore individual variables with dplyr
## 2.1 Possible values/ranges of each variable
```{r}
gapminder %>%
  filter (country == "Canada") %>%
  ggplot(aes(year, lifeExp)) +
  geom_point () 
  
```

## 2.2 Tailor to the variable at hand
1. Canada's life expense
```{r}
gapminder %>%
  filter (country == "Canada") %>%
  summarise("min_lifeExp" = min(lifeExp), "mean_lifeExp" = mean(lifeExp), "max_lifeExp" = max(lifeExp))

```

2. Asia's population
```{r}
gapminder %>%
  filter (continent == "Asia") %>%
  ggplot (aes(year, pop)) + 
  geom_point()
```


# Exercise 3: Explore various plot types
## 3.1 A scatterplot of two quantitative variables.
```{r}
gapminder %>%
  group_by (continent) %>%
  ggplot (aes(pop, lifeExp, continent)) + 
  geom_point(alpha = 0.2) +
  scale_x_log10() 
```


## 3.2 one other plot besides a scatterplot
```{r}
gapminder %>%
  filter (country == "Canada" | country == "China") %>%
  ggplot (aes(x=year, y=lifeExp, col=country)) + 
  geom_point()
```

  

# Recycling 
```{r}
filter(gapminder, country == c("Rwanda", "Afghanistan"))
```

The analyst’s intent was to get the data for Rwanda and Afghanistan. To correct it, an "or" operator should be used
```{r}
filter(gapminder, country == "Rwanda" | country == "Afghanistan")

```
