---
title: "hw03_dplyr_ggplot2"
author: "Xuemeng Li"
date: "29/09/2019"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview
The goal is to manipulate and explore a dataset with the dplyr package, complemented by visualizations made with ggplot2. 

```{r}
library(gapminder)
library(tidyverse)
library(gridExtra)
```


# Task Option 2
Get the maximum and minimum of GDP per capita for all continents.

## 1. A table to show 
```{r}
gapminder %>% 
  select (gdpPercap, continent) %>%
  group_by (continent) %>%
  summarize(max_cap = max(gdpPercap), min_cap = min(gdpPercap), median_cap = median(gdpPercap))
```


## 2. Visulize the data in a figure to get better comparison


```{r}
m <- gapminder %>% 
  select (gdpPercap, continent) %>%
  group_by (continent) %>%
  summarize(max_cap = max(gdpPercap), min_cap = min(gdpPercap), median_cap = median(gdpPercap))

ggplot(gapminder, aes(continent,gdpPercap)) +
  scale_y_log10("gdpPercap(log(usd))")+
  geom_point(data = m, aes(x = continent, y = max_cap), colour = "blue", size = 3) +
  geom_point(data = m, aes(x = continent, y = min_cap), colour = "red", size = 3) +
  theme_classic()
```

```{r}
gapminder %>% 
  select (gdpPercap, continent) %>%
  group_by (continent) %>%
  ggplot (aes (continent,gdpPercap)) +
  geom_boxplot() +
  scale_y_log10("gdpPercap(log(usd))") +
  theme_classic()
```

# Task Option 3
Look at the spread of GDP per capita within the continents.

## Point spreading
```{r}
ggplot(gapminder, aes(x = continent, y = gdpPercap)) +
  scale_y_log10() +
  geom_jitter(position = position_jitter(width = 0.1, height = 0), alpha = 1/4)
```


# Task Option 4
Compute a trimmed mean of life expectancy for different years. Or a weighted mean, weighting by population. Just try something other than the plain vanilla mean.

## Compare the mean for each continent with the trimmed mean 
```{r}
p1<-gapminder %>% 
  select (lifeExp, year, continent) %>%
  group_by(continent, year) %>%
  mutate (mean_lifeExp = mean(lifeExp)) %>%
  mutate (trimmed_mean_lifeExp = mean(lifeExp, trim = 0.1)) %>%
  select (year, continent, mean_lifeExp, trimmed_mean_lifeExp) %>%
  ggplot(aes(x = year, y = mean_lifeExp, col = continent))+
  geom_point (size = 1) +
  geom_line (size = 0.5) 

p2<-gapminder %>% 
  select (lifeExp, year, continent) %>%
  group_by(continent, year) %>%
  mutate (mean_lifeExp = mean(lifeExp)) %>%
  mutate (trimmed_mean_lifeExp = mean(lifeExp, trim = 0.1)) %>%
  select (year, continent, mean_lifeExp, trimmed_mean_lifeExp) %>%
  ggplot(aes(x = year, y = trimmed_mean_lifeExp, col = continent))+
  geom_point (size = 1) +
  geom_line (size = 0.5) 

grid.arrange(p1, p2, nrow = 2)
```

# Task Option 5
How is life expectancy changing over time on different continents?

## The mean of life expectancy of each continents are calculated and compared as follow:
```{r}
gapminder %>% 
  select (lifeExp, year, continent) %>%
  group_by(continent, year) %>%
  mutate (mean_lifeExp = mean(lifeExp)) %>%
  select (mean_lifeExp, year, continent) %>%
  ggplot(aes(x = year, y = mean_lifeExp, col = continent)) +
  geom_point (size = 1) +
  geom_line (size = 0.5) 
```


