---
title: "hw01_gapminder"
author: "Xuemeng Li"
date: "24/09/2019"
output: 
  html_document:
    keep_md: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This is exercise 2 in homework 1. In this exercise, we will explore gapminder dataset

## Basic Usage

* Source the library
```{r}
library(gapminder)
library(dplyr)

```

* Get intro about the dataset from the R Studio: 

```{r eval = FALSE, comment = NA}
?gapminder
```
* Here are the headers for gapminder
```{r}
head(gapminder)
```

* Here is a Summary 
```{r}
summary(gapminder)
```


* Here is a graph for the average life exp for Combodia from 1950 to 2000
```{r}
plot(lifeExp ~ year, gapminder, subset = country == "Cambodia", type = "b")
```

## Analysis with dataset

* How many unique countries does the data contain, by continent?
```{r}
gapminder %>%
  group_by(continent) %>%
  summarize(n_obs = n(), n_countries = n_distinct(country))
```


*By continent, which country experienced the sharpest 5-year drop in life expectancy and what was the drop?

```{r}
gapminder %>%
  group_by(continent, country) %>%
  select(country, year, continent, lifeExp) %>%
  mutate(le_delta = lifeExp - lag(lifeExp)) %>%
  summarize(worst_le_delta = min(le_delta, na.rm = TRUE)) %>%
  filter(min_rank(worst_le_delta) < 2) %>%
  arrange(worst_le_delta)

```

